Imperative Programming with Variational Effects

### 2018-08-03 paper/paper.tex
work on listings

\usepackage{setspace} 
\singlespacing
\usepackage{graphicx}                                        
\usepackage{amssymb}                                         
\usepackage{amsmath}                                         
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{algorithmic}
\usepackage{url}
\usepackage{tikz}         
\usetikzlibrary{matrix}
\usepackage{listings}
\usepackage{tabu}
\usepackage{mathptmx}
\usepackage{fullpage}
\usepackage{enumerate}
\usepackage{fancyvrb}
\usepackage{mathpartir}
\usepackage{lambda}
\usepackage{cc}
\usepackage{cc-ext}
\usepackage{alltt}

@@

@@

\begin{hprogram}
type Dim = String
data Cond =
    Lit Bool
  | Ref Dim
  | Not Cond
  | And Cond Cond
  | Or Cond Cond

data V a = One a | Chc Cond (V a) (V a) 
\end{hprogram}

@@

\begin{hprogram}
instance Functor V where
  fmap f (One a) = One (f a)
  fmap f (Chc d l r) = Chc d (fmap f l) (fmap f r)

@@

instance Monad V where
  return = One
  (One v) >>= f = f v
  (Chc d l r) >>= f = Chc d (l >>= f) (r >>= f)
\end{hprogram}

@@

\begin{hprogram}
data Arith = N Int | Add Arith Arith | Mul Arith Arith
\end{hprogram}

@@

\begin{hprogram}
eval :: Arith -> Int
eval (N i) = i
eval (Add l r) = eval l + eval r
eval (Mul l r) = eval l * eval r
\end{hprogram}

@@

\begin{hprogram}
data Arith = N Int | Add Arith Arith | Mul Arith Arith | AChc Cond Arith Arith
\end{hprogram}

@@

\begin{hprogram}
eval :: Arith -> V Int
eval (N i) = pure i
eval (Add l r) = (+) <$> eval l <*> eval r
eval (Mul l r) = (*) <$> eval l <*> eval r
eval (AChc d l r) = Chc d (eval l) (eval r)
\end{hprogram}

@@

\begin{figure}
\begin{syntax}
\text{(Arithmetic expressions)}\\
a &::=& n & \textit{Integer literal} \\
& | & x & \textit{Variable reference} \\
& | & a + a & \textit{Addition} \\
\text{(Boolean literals)} \\
b &::=& \CCkeyw{true} \\
& | & \CCkeyw{false} \\
\text{(Boolean expressions)} \\
e &::=& b & \textit{Boolean literal} \\
& | & \CCkeyw{not}\ e & \textit{Negation} \\
& | & e\ \CCkeyw{and}\ e & \textit{Conjunction} \\
& | & a < a & \textit{Less} \\
\text{(Statements)} \\
s &::=& \CCkeyw{skip} & \textit{Noop} \\
& | & x \coloneqq a & \textit{Assignment} \\
& | & s\ ;\ s & \textit{Sequencing} \\
& | & \CCkeyw{if}\ e\ \CCkeyw{then}\ s\ \CCkeyw{else}\ s & \textit{Conditional} \\
& | & \CCkeyw{while}\ e\ \CCkeyw{do}\ s & \textit{Looping} \\
& | & \CCkeyw{throw}\ a & \textit{Throw error} \\
& | & \CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{in}\ s & \textit{Catch error}
\end{syntax}
\caption{Syntax of IMP with Exceptions}
\label{fig:impsyn}
\end{figure}

We begin by establishing the syntax and semantics of plain IMP before turning to variational IMP. The syntax of plain IMP is defined in Figure \ref{fig:impsyn}.
At its base IMP defines simple arithmetic expressions made up of integer literals, variable references, and addition. IMP also supports boolean expressions
made up of literals, negation, conjunction, and less-than comparison of arithmetic expressions. Statements include variable assignment, sequencing, and
control structures. Only arithmetic expressions are allowed in variable assignments, and only boolean expressions are allowed in the conditions of control
structures, thus obviating the need for a type system. Finally, we have extended the definition of the language to include basic support for catchable exceptions
as one of the effects that we wish to formally define for variational execution. \CCkeyw{throw} statements take an arithmetic expression and produce an
error with the value the expression evaluates to. \CCkeyw{catch} statements allow for the handling of any thrown error values in their main block via a
handler block.

We now define a formal big-step semantics for this non-variational version of IMP, which we can then use as a point of reference for our variational semantics.
Figure \ref{fig:impexpr} defines the big-step semantics of arithmetic and boolean expressions. Expressions are evaluated in the context of the current variable
store and are otherwise very straightforward in their definition. Figure \ref{fig:impstmt} defines the big-step semantics of IMP statements.

@@

\def \nothing {\bullet}
\def \BigErr {\infer [S-Err] { } {(n,S,s) \Downarrow_S (n,S)}}
\def \BigSkip {\infer [S-Skip] { } {(E,S,\CCkeyw{skip}) \Downarrow_S (E,S)}}
\def \BigAssn {\infer [S-Assn] {(S,a) \Downarrow_A n} {(\nothing,S,x \coloneqq a) \Downarrow_S (\nothing,S[x \rightarrow n])}}
\def \BigSeq {\infer [S-Seq] {(E,S,s) \Downarrow_S (E',S') \\ (E',S',s') \Downarrow_S (E'',S'')} {(E,S,s\ ;\ s') \Downarrow_S (E'',S'')}}
\def \BigIfT {\infer [S-IfT] {(S,e) \Downarrow_B \CCkeyw{true} \\ (\nothing,S,s) \Downarrow_S (E,S')} {(\nothing,S,\CCkeyw{if}\ e\ \CCkeyw{then}\ s\ \CCkeyw{else}\ s') \Downarrow_S (E,S')}}
\def \BigIfF {\infer [S-IfF] {(S,e) \Downarrow_B \CCkeyw{false} \\ (\nothing,S,s') \Downarrow_S (E,S')} {(\nothing,S,\CCkeyw{if}\ e\ \CCkeyw{then}\ s\ \CCkeyw{else}\ s') \Downarrow_S (E,S')}}
\def \BigWhileT {\infer [S-WhileT] {(S,e) \Downarrow_B \CCkeyw{true} \\ (\nothing,S,s) \Downarrow_S (E,S') \\ (E,S',\CCkeyw{while}\ e\ \CCkeyw{do}\ s) \Downarrow_S (E',S'')} {(\nothing,S,\CCkeyw{while}\ e\ \CCkeyw{do}\ s) \Downarrow_S (E',S'')}}
\def \BigWhileF {\infer [S-WhileF] {(S,e) \Downarrow_B \CCkeyw{false}} {(E,S,\CCkeyw{while}\ e\ \CCkeyw{do}\ s) \Downarrow_S (E,S)}}
\def \BigThrow {\infer [S-Throw] {(S,a) \Downarrow_A n} {(\nothing,S,\CCkeyw{throw}\ a) \Downarrow_S (n,S)}}
\def \BigTry {\infer [S-Try] {(\nothing,S,s) \Downarrow_S (\nothing,S')} {(\nothing,S,\CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{in}\ s') \Downarrow_S (\nothing,S')}}
\def \BigCatch {\infer [S-Catch] {(\nothing,S,s) \Downarrow_S (n,S') \\ (\nothing,S'[x \rightarrow n],s') \Downarrow_S (E,S'')} {(\nothing,S,\CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{in}\ s') \Downarrow_S (E,S''[x \rightarrow S'(x)])}}

@@

\begin{figure}
\begin{syntax}
\text{(Boolean literals)} \\
b &::=& \CCkeyw{true} \\
& | & \CCkeyw{false} \\
\text{(Conditions)}\\
C &::=& b & \textit{Boolean literal} \\
& | & D & \textit{Dimension name} \\
& | & \neg C & \textit{Negation} \\
& | & C \vee C & \textit{Disjunction} \\
& | & C \wedge C & \textit{Conjunction} \\
\text{(Arithmetic expressions)}\\
a &::=& n & \textit{Integer literal} \\
& | & x & \textit{Variable reference} \\
& | & a + a & \textit{Addition} \\
& | & \tagtree{C}{a}{a} & \textit{Choice} \\
\text{(Boolean expressions)} \\
e &::=& b & \textit{Boolean literal} \\
& | & \CCkeyw{not}\ e & \textit{Negation} \\
& | & e\ \CCkeyw{and}\ e & \textit{Conjunction} \\
& | & a < a & \textit{Less} \\
& | & \tagtree{C}{e}{e} & \textit{Choice} \\
\text{(Statements)} \\
s &::=& \CCkeyw{skip} & \textit{Noop} \\
& | & x \coloneqq a & \textit{Assignment} \\
& | & s\ ;\ s & \textit{Sequencing} \\
& | & \CCkeyw{if}\ e\ \CCkeyw{then}\ s\ \CCkeyw{else}\ s & \textit{Conditional} \\
& | & \CCkeyw{while}\ e\ \CCkeyw{do}\ s & \textit{Looping} \\
& | & \CCkeyw{throw}\ a & \textit{Throw error} \\
& | & \CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{in}\ s & \textit{Catch error} \\
& | & \tagtree{C}{s}{s} & \textit{Choice}
\end{syntax}
\caption{Syntax of Variational IMP with Effects}
\label{fig:vimpsyn}
\end{figure}

@@

\newcommand{\GetCtx}[2]{\mathit{ctx}(#1,#2)}
\def \gc {\GetCtx{C}{E}}
\def \T {\mathit{true}}
\def \F {\mathit{false}}
\def \BigVUnsat {\infer [V-Unsat] {\mathit{unsat}(\gc)} {(C,S,E,s) \Downarrow_{V} (S,E)}}
\def \BigVSat {\infer [V-Sat] {\mathit{sat}(\gc) \\ (C,S,E,s) \Downarrow_{VS} (S',E')} {(C,S,E,s) \Downarrow_{V} (S',E')}}
\def \BigVSkip {\infer[VS-Skip] { } {(C,S,E,\CCkeyw{skip}) \Downarrow_{VS} (S,E)}}
\def \BigVAssn {\infer[VS-Assn] {(\gc,S,a) \Downarrow_{VA} u} {(C,S,E,x \coloneqq a) \Downarrow_{VS} (S[(\gc,x) \rightarrow u],E)}}
\def \BigVSeq {\infer [VS-Seq] {(C,S,E,s) \Downarrow_{V} (S',E') \\ (C',S',E',s') \Downarrow_{V} (S'',E'')} {(C,S,E,s\ ;\ s') \Downarrow_{VS} (S'',E'')}}
\def \BigVIf {\infer [VS-If] {(\gc,S,e) \Downarrow_{VB} v \\ (C \wedge \mathit{whenTrue}(v),S,E,s) \Downarrow_{V} (S',E') \\ (C \wedge \mathit{whenFalse}(v),S',E') \Downarrow_{V} (S'',E'')} {(C,S,E,\CCkeyw{if}\ e\ \CCkeyw{then}\ s\ \CCkeyw{else}\ s') \Downarrow_{VS} (S'',E'')}}
\def \BigVWhile {\infer [VS-While] {(\GetCtx{C}{E},S,e) \Downarrow_{VB} v \\ (C \wedge \mathit{whenTrue}(v),S,E,s) \Downarrow_{V} (S',E') \\ (C \wedge \mathit{whenTrue}(v),S',E',\CCkeyw{while}\ e\ \CCkeyw{do}\ s) \Downarrow_{V} (S'',E'')} {(C,S,E,\CCkeyw{while}\ e\ \CCkeyw{do}\ s) \Downarrow_{VS} (S'',E'')}}
\def \BigVThrow {\infer [VS-Throw] {(\GetCtx{C}{l:E},S,a) \Downarrow_{VA} u} {(C,S,l:E,\CCkeyw{throw}\ a) \Downarrow_{VS} (S,((C,u):l):E)}}
\def \BigVTry {\infer [VS-Try] {(C,S,[]:E,s) \Downarrow_{V} (S',[]:E)} {(C,S,E,\CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{in}\ s') \Downarrow_{VS} (S',E)}}
\def \BigVCatch {\infer [VS-Catch] {(C,S,[]:E,s) \Downarrow_{V} (S',l:E) \\ (\GetCtx{C}{(l:[])},S'[(\GetCtx{C}{E},x) \rightarrow \mathit{combine}(l)],E,s') \Downarrow_{V} (S'',E')} {(C,S,E,\CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{in}\ s') \Downarrow_{VS} (S''[(\CCkeyw{true},x) \rightarrow S'(\CCkeyw{true},x)],E')}}
\def \BigVChc {\infer [VS-Chc] {(C \wedge C',S,E,s) \Downarrow_{V} (S',E') \\ (C \wedge \neg C',S',E',s') \Downarrow_{V} (S'',E'')} {(C,S,E,\tagtree{C'}{s}{s'}) \Downarrow_{VS} (S'',E'')}}

>>>>

\usepackage{setspace} 
\singlespacing
\usepackage{graphicx}                                        
\usepackage{amssymb}                                         
\usepackage{amsmath}                                         
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{algorithmic}
\usepackage{url}
\usepackage{tikz}         
\usetikzlibrary{matrix}
\usepackage{listings}
\usepackage{tabu}
\usepackage{mathptmx}
\usepackage{fullpage}
\usepackage{enumerate}
\usepackage{fancyvrb}
\usepackage{mathpartir}
\usepackage{lambda}
\usepackage{cc}
\usepackage{cc-ext}
\usepackage{alltt}
\usepackage{textcomp}

@@

\usepackage{listings}

\lstset{columns=fullflexible,
        upquote=true,
        mathescape=true,
        literate=
               {=>}{$\Rightarrow{}$}{1}
               {.}{$\circ{}\ $}{1}
               {<*>}{$<\!\!\!\!\!*\!\!\!\!\!>{}$}{1}
               {>>=}{$>\!\!\!>\!={}$}{1}
               {->}{$\rightarrow{}$}{1}
               {<>}{$<\!\!\!\$\!\!\!>{}$}{1}
               {'}{$^\prime{}$}{1},
         morekeywords={data,instance,where,type,if,then,else,otherwise}
        }

@@

\begin{lstlisting}
type Dim = String
data Cond =
    Lit Bool
  | Ref Dim
  | Not Cond
  | Cond :&&: Cond
  | Cond :||: Cond

data V a = One a | Chc Cond (V a) (V a)
\end{lstlisting}

@@

\begin{lstlisting}
instance Functor V where
  fmap f (One a) = One (f a)
  fmap f (Chc d l r) = Chc d (fmap f l) (fmap f r)

@@

instance Monad V where
  return = One
  (One v) >>= f = f v
  (Chc d l r) >>= f = Chc d (l >>= f) (r >>= f)
\end{lstlisting}

@@

\begin{lstlisting}
data Arith = N Int | Add Arith Arith | Mul Arith Arith
\end{lstlisting}

@@

\begin{lstlisting}
eval :: Arith -> Int
eval (N i) = i
eval (Add l r) = eval l + eval r
eval (Mul l r) = eval l * eval r
\end{lstlisting}

@@

\begin{lstlisting}
data Arith = N Int | Add Arith Arith | Mul Arith Arith | AChc Cond Arith Arith
\end{lstlisting}

@@

\begin{lstlisting}
eval :: Arith -> V Int
eval (N i) = pure i
eval (Add l r) = (+) <> eval l <*> eval r
eval (Mul l r) = (*) <> eval l <*> eval r
eval (AChc d l r) = Chc d (eval l) (eval r)
\end{lstlisting}

@@

\begin{figure}
\begin{syntax}
\text{(Arithmetic expressions)}\\
a &::=& n & \textit{Integer literal} \\
& | & x & \textit{Variable reference} \\
& | & a + a & \textit{Addition} \\
\text{(Boolean literals)} \\
b &::=& \CCkeyw{true} \\
& | & \CCkeyw{false} \\
\text{(Boolean expressions)} \\
e &::=& b & \textit{Boolean literal} \\
& | & \CCkeyw{not}\ e & \textit{Negation} \\
& | & e\ \CCkeyw{and}\ e & \textit{Conjunction} \\
& | & a < a & \textit{Less} \\
\text{(Statements)} \\
s &::=& \CCkeyw{skip} & \textit{Noop} \\
& | & x \coloneqq a & \textit{Assignment} \\
& | & s\ ;\ s & \textit{Sequencing} \\
& | & \CCkeyw{if}\ e\ \CCkeyw{then}\ s\ \CCkeyw{else}\ s & \textit{Conditional} \\
& | & \CCkeyw{while}\ e\ \CCkeyw{do}\ s & \textit{Looping} \\
& | & \CCkeyw{throw}\ a & \textit{Throw error} \\
& | & \CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{with}\ s & \textit{Catch error}
\end{syntax}
\caption{Syntax of IMP with Exceptions}
\label{fig:impsyn}
\end{figure}

We begin by establishing the syntax and semantics of plain IMP before turning to variational IMP. The syntax of plain IMP is defined in Figure \ref{fig:impsyn}.
At its base IMP defines simple arithmetic expressions made up of integer literals, variable references, and addition. IMP also supports boolean expressions
made up of literals, negation, conjunction, and less-than comparison of arithmetic expressions. Statements include variable assignment, sequencing, and
control structures. Only arithmetic expressions are allowed in variable assignments, and only boolean expressions are allowed in the conditions of control
structures, thus obviating the need for a type system. Finally, we have extended the definition of the language to include basic support for catchable exceptions
as one of the effects that we wish to formally define for variational execution. \CCkeyw{throw} statements take an arithmetic expression and produce an
error with the value the expression evaluates to. \CCkeyw{try \ldots catch} statements allow for the handling of any thrown error values in their main block via a
handler block.

We now define a formal big-step semantics for this non-variational version of IMP, which we can then use as a point of reference for our variational semantics.
Figure \ref{fig:impexpr} defines the big-step semantics of arithmetic and boolean expressions. Expressions are evaluated in the context of the current variable
store and are otherwise straightforward in their definition. Figure \ref{fig:impstmt} defines the big-step semantics of IMP statements. Statements have two possible
side effects: updating the variable store or updating the current error context.

@@

\def \nothing {\bullet}
\def \BigErr {\infer [S-Err] { } {(n,S,s) \Downarrow_S (n,S)}}
\def \BigSkip {\infer [S-Skip] { } {(E,S,\CCkeyw{skip}) \Downarrow_S (E,S)}}
\def \BigAssn {\infer [S-Assn] {(S,a) \Downarrow_A n} {(\nothing,S,x \coloneqq a) \Downarrow_S (\nothing,S[x \rightarrow n])}}
\def \BigSeq {\infer [S-Seq] {(E,S,s) \Downarrow_S (E',S') \\ (E',S',s') \Downarrow_S (E'',S'')} {(E,S,s\ ;\ s') \Downarrow_S (E'',S'')}}
\def \BigIfT {\infer [S-IfT] {(S,e) \Downarrow_B \CCkeyw{true} \\ (\nothing,S,s) \Downarrow_S (E,S')} {(\nothing,S,\CCkeyw{if}\ e\ \CCkeyw{then}\ s\ \CCkeyw{else}\ s') \Downarrow_S (E,S')}}
\def \BigIfF {\infer [S-IfF] {(S,e) \Downarrow_B \CCkeyw{false} \\ (\nothing,S,s') \Downarrow_S (E,S')} {(\nothing,S,\CCkeyw{if}\ e\ \CCkeyw{then}\ s\ \CCkeyw{else}\ s') \Downarrow_S (E,S')}}
\def \BigWhileT {\infer [S-WhileT] {(S,e) \Downarrow_B \CCkeyw{true} \\ (\nothing,S,s) \Downarrow_S (E,S') \\ (E,S',\CCkeyw{while}\ e\ \CCkeyw{do}\ s) \Downarrow_S (E',S'')} {(\nothing,S,\CCkeyw{while}\ e\ \CCkeyw{do}\ s) \Downarrow_S (E',S'')}}
\def \BigWhileF {\infer [S-WhileF] {(S,e) \Downarrow_B \CCkeyw{false}} {(E,S,\CCkeyw{while}\ e\ \CCkeyw{do}\ s) \Downarrow_S (E,S)}}
\def \BigThrow {\infer [S-Throw] {(S,a) \Downarrow_A n} {(\nothing,S,\CCkeyw{throw}\ a) \Downarrow_S (n,S)}}
\def \BigTry {\infer [S-Try] {(\nothing,S,s) \Downarrow_S (\nothing,S')} {(\nothing,S,\CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{with}\ s') \Downarrow_S (\nothing,S')}}
\def \BigCatch {\infer [S-Catch] {(\nothing,S,s) \Downarrow_S (n,S') \\ (\nothing,S'[x \rightarrow n],s') \Downarrow_S (E,S'')} {(\nothing,S,\CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{with}\ s') \Downarrow_S (E,S'')}}

@@

\subsection{Variational Operations}

In order to successfully relate variational IMP back to its non-variational counterpart
we must define configuration and selection for VIMP. We assume solving for satisfiability
as a primitive operation; in practice this can be implemented by offloading these calls to
a SAT solver. We begin by defining the general case of selection and configuration for choice
trees and defer discussion of the implementation of these operations for the variational store to its own section.

\begin{lstlisting}
unsat = not . sat
taut = unsat . Not
x => y = taut ((Not x) :||: y)

sel _ (One x) = x
sel c (Chc c' l r)
  | c => c' = sel c l
  | c => (Not c') = sel c r
  | otherwise = Chc c' (sel c l) (sel c r)

evalCond ds (Lit b) = b
evalCond ds (Ref x) = x `elem` ds
evalCond ds (Not c) = not (evalCond c)
evalCond ds (And c c') = evalCond c && evalCond c'
evalCond ds (Or c c') = evalCond c || evalCond c'

conf _ (One x) = x
conf ds (Chc c l r) = if evalCond ds c then conf ds l else conf ds r
\end{lstlisting}

\begin{figure}
\begin{syntax}
\text{(Boolean literals)} \\
b &::=& \CCkeyw{true} \\
& | & \CCkeyw{false} \\
\text{(Conditions)}\\
C &::=& b & \textit{Boolean literal} \\
& | & D & \textit{Dimension name} \\
& | & \neg C & \textit{Negation} \\
& | & C \vee C & \textit{Disjunction} \\
& | & C \wedge C & \textit{Conjunction} \\
\text{(Arithmetic expressions)}\\
a &::=& n & \textit{Integer literal} \\
& | & x & \textit{Variable reference} \\
& | & a + a & \textit{Addition} \\
& | & \tagtree{C}{a}{a} & \textit{Choice} \\
\text{(Boolean expressions)} \\
e &::=& b & \textit{Boolean literal} \\
& | & \CCkeyw{not}\ e & \textit{Negation} \\
& | & e\ \CCkeyw{and}\ e & \textit{Conjunction} \\
& | & a < a & \textit{Less} \\
& | & \tagtree{C}{e}{e} & \textit{Choice} \\
\text{(Statements)} \\
s &::=& \CCkeyw{skip} & \textit{Noop} \\
& | & x \coloneqq a & \textit{Assignment} \\
& | & s\ ;\ s & \textit{Sequencing} \\
& | & \CCkeyw{if}\ e\ \CCkeyw{then}\ s\ \CCkeyw{else}\ s & \textit{Conditional} \\
& | & \CCkeyw{while}\ e\ \CCkeyw{do}\ s & \textit{Looping} \\
& | & \CCkeyw{throw}\ a & \textit{Throw error} \\
& | & \CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{with}\ s & \textit{Catch error} \\
& | & \tagtree{C}{s}{s} & \textit{Choice}
\end{syntax}
\caption{Syntax of Variational IMP with Effects}
\label{fig:vimpsyn}
\end{figure}

@@

\newcommand{\GetCtx}[2]{\mathit{ctx}(#1,#2)}
\def \gc {\GetCtx{C}{E}}
\def \T {\mathit{true}}
\def \F {\mathit{false}}
\def \BigVUnsat {\infer [V-Unsat] {\mathit{unsat}(\gc)} {(C,S,E,s) \Downarrow_{V} (S,E)}}
\def \BigVSat {\infer [V-Sat] {\mathit{sat}(\gc) \\ (C,S,E,s) \Downarrow_{VS} (S',E')} {(C,S,E,s) \Downarrow_{V} (S',E')}}
\def \BigVSkip {\infer[VS-Skip] { } {(C,S,E,\CCkeyw{skip}) \Downarrow_{VS} (S,E)}}
\def \BigVAssn {\infer[VS-Assn] {(\gc,S,a) \Downarrow_{VA} u} {(C,S,E,x \coloneqq a) \Downarrow_{VS} (S[(\gc,x) \rightarrow u],E)}}
\def \BigVSeq {\infer [VS-Seq] {(C,S,E,s) \Downarrow_{V} (S',E') \\ (C',S',E',s') \Downarrow_{V} (S'',E'')} {(C,S,E,s\ ;\ s') \Downarrow_{VS} (S'',E'')}}
\def \BigVIf {\infer [VS-If] {(\gc,S,e) \Downarrow_{VB} v \\ (C \wedge \mathit{whenTrue}(v),S,E,s) \Downarrow_{V} (S',E') \\ (C \wedge \mathit{whenFalse}(v),S',E') \Downarrow_{V} (S'',E'')} {(C,S,E,\CCkeyw{if}\ e\ \CCkeyw{then}\ s\ \CCkeyw{else}\ s') \Downarrow_{VS} (S'',E'')}}
\def \BigVWhile {\infer [VS-While] {(\GetCtx{C}{E},S,e) \Downarrow_{VB} v \\ (C \wedge \mathit{whenTrue}(v),S,E,s) \Downarrow_{V} (S',E') \\ (C \wedge \mathit{whenTrue}(v),S',E',\CCkeyw{while}\ e\ \CCkeyw{do}\ s) \Downarrow_{V} (S'',E'')} {(C,S,E,\CCkeyw{while}\ e\ \CCkeyw{do}\ s) \Downarrow_{VS} (S'',E'')}}
\def \BigVThrow {\infer [VS-Throw] {(\GetCtx{C}{l:E},S,a) \Downarrow_{VA} u} {(C,S,l:E,\CCkeyw{throw}\ a) \Downarrow_{VS} (S,((C,u):l):E)}}
\def \BigVTry {\infer [VS-Try] {(C,S,[]:E,s) \Downarrow_{V} (S',[]:E)} {(C,S,E,\CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{with}\ s') \Downarrow_{VS} (S',E)}}
\def \BigVCatch {\infer [VS-Catch] {(C,S,[]:E,s) \Downarrow_{V} (S',l:E) \\ (\GetCtx{C}{(l:[])},S'[(\GetCtx{C}{E},x) \rightarrow \mathit{combine}(l)],E,s') \Downarrow_{V} (S'',E')} {(C,S,E,\CCkeyw{try}\ s\ \CCkeyw{catch}\ x\ \CCkeyw{with}\ s') \Downarrow_{VS} (S'',E')}}
\def \BigVChc {\infer [VS-Chc] {(C \wedge C',S,E,s) \Downarrow_{V} (S',E') \\ (C \wedge \neg C',S',E',s') \Downarrow_{V} (S'',E'')} {(C,S,E,\tagtree{C'}{s}{s'}) \Downarrow_{VS} (S'',E'')}}
